%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
%%% Short-Term Forecasting in SIR Systems:  
%%% Last edited by Ravi (June 9, 2015).  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
  
\documentclass[11pt]{article} 
%\usepackage{latex8}  
\usepackage{times}  
%\usepackage{epsfig}  
%\usepackage{psfig}  
\usepackage{amsfonts}  
\usepackage{latexsym}  
\usepackage{amsmath}  
\usepackage[latin1]{inputenc}  
\usepackage{varioref}  
\usepackage{cite}  
  
\usepackage{fancybox}  
%% \usepackage{shadow}  
\usepackage{times}  
\usepackage{amsmath}  
\usepackage{amsfonts}  
\usepackage{latexsym}  
\usepackage{color}  
\usepackage{graphicx}  
\usepackage{subfig}  
\usepackage{lscape}  
\usepackage{url}

\usepackage{epigraph}  
%%\usepackage{patchcmd}  

%%\usepackage{framed}  

  
\setlength{\textheight}{9 in}  
\setlength{\textwidth}{6.3 in}  
\setlength{\oddsidemargin}{-0.0125 in}  
\setlength{\evensidemargin}{-0.0125 in}  
\setlength{\topmargin}{-0.5 in}  
\setlength{\parskip}{2pt}  

\setlength{\itemsep}{1pt}  

%% For use with the epigraph package

\setlength\epigraphwidth{10cm}
\setlength\epigraphrule{0pt}

%%\makeatletter
%%\patchcommand{\epigraph}{\@epitext{#1}}{\itshape\@epitext{#1}}{}{}
%%\makeatother
   
  
\newtheorem{theorem}{Theorem}[section]  
\newtheorem{lemma}[theorem]{Lemma}  
\newtheorem{corollary}[theorem]{Corollary}  
\newtheorem{fact}[theorem]{Fact}  
\newtheorem{claim}[theorem]{Claim}  
\newtheorem{observation}[theorem]{Observation}  
\newtheorem{definition}[theorem]{Definition}  
\newtheorem{proposition}[theorem]{Proposition}  
\newtheorem{example}[theorem]{Example}  
  
\newcommand{\calb}{\mbox{${\cal B}$}}  
\newcommand{\calc}{\mbox{${\cal C}$}}  
\newcommand{\calcp}{\mbox{${\cal C}'$}}  

\newcommand{\calco}{\mbox{$\mathcal{C}_1$}}
\newcommand{\calct}{\mbox{$\mathcal{C}_2$}}
\newcommand{\calcop}{\mbox{$\mathcal{C}_1'$}}
\newcommand{\calcodp}{\mbox{$\mathcal{C}_1''$}}

\newcommand{\calcdp}{\mbox{${\cal C}''$}}  
\newcommand{\calf}{\mbox{${\cal F}$}}  
\newcommand{\cale}{\mbox{${\cal E}$}}  
\newcommand{\cali}{\mbox{${\cal I}$}}  
\newcommand{\calp}{\mbox{${\cal P}$}}  
\newcommand{\cals}{\mbox{${\cal S}$}}  
\newcommand{\calt}{\mbox{${\cal T}$}}  
  
\newcommand{\calh}{\mbox{${\cal H}$}}  
\newcommand{\calr}{\mbox{${\cal R}$}}  
\newcommand{\cald}{\mbox{${\cal D}$}}  


\newcommand{\cnp}{\mbox{\textbf{NP}}}
\newcommand{\cnump}{\mbox{\textbf{\#P}}}

\newcommand{\ttrue}{\textsc{True}}
\newcommand{\tfalse}{\textsc{False}}
  
%% \newcommand{\reach}{{\textsc{Reachability}}}
%% \newcommand{\treach}{{$t$-\textsc{Reachability}}}

% box for end of proof  
\newcommand{\QED}{\hfill\rule{2mm}{2mm}\medskip}   

%% Additional newcommands from Anil.

\newcommand{\opt}{\text{OPT}}
\newcommand{\eopt}{E_{\text{OPT}}}
\newcommand{\prob}{\textsc{Spectral Radius Minimization}}
\newcommand{\expect}{\mathrm{Exp}}
\DeclareMathOperator*{\nodes}{nodes}
\DeclareMathOperator*{\walks}{walks}
\DeclareMathOperator*{\ct}{count}
  
  

\begin{document} 
\include{titlepage}
%%\maketitle  

\begin{center}
\Large{\textbf{Algorithmic theory of Forecasting 
               Contagion Dynamics} \\ \smallskip
       \textbf{over Networks -- Part~I~:~ General Graphs}}
\end{center}

\medskip

\thispagestyle{empty}

\begin{center}
\textsc{Madhav V. Marathe}$^1$ \hspace{0.2in}
\textsc{S. S. Ravi}$^{2,3}$ \hspace{0.2in}
\textsc{Daniel J. Rosenkrantz}$^3$ \\ 
\vspace*{1ex}
\textsc{Richard E. Stearns}$^3$ \hspace{0.2in}
\textsc{Anil K. Vullikanti}$^1$ 
\end{center}

\setcounter{footnote}{1}
\footnotetext[1]{
Network Dynamics and Simulation Sciences Laboratory,
Biocomplexity Institute and 
Department of Computer Science, Virginia Tech,
Blacksburg, VA 24061.
\textsf{Email:} \texttt{\{mmarathe, ssravi, akumar\}@vbi.vt.edu}. \smallskip
}

\setcounter{footnote}{2}
\footnotetext[2]{
Network Dynamics and Simulation Sciences Laboratory,
Biocomplexity Institute, Virginia Tech,
Blacksburg, VA 24061.
\textsf{Email:} \texttt{ssravi@vbi.vt.edu}. \smallskip
}

\setcounter{footnote}{3}
\footnotetext[3]{
Department of Computer Science,
University at Albany -- State University of New York,
Albany, NY 12222. \newline
\textsf{Email:} \texttt{drosenkrantz@gmail.com} ~and~ 
                \texttt{thestearns2@gmail.com}
}

\begin{center}
\textbf{Abstract}
\end{center}

%%\begin{center}
%%\fbox{\textbf{To be revised.}}
%%\end{center}

The Susceptible-Infected-Recovered (SIR) model of disease
propagation is widely used in epidemiology.  
Under this model, we study short-term forecasting problems
where the goal is to predict a given system's 
behavior a few time steps into the future. 
An example of such a problem is the following: find
the probability that there will be at least $q$ new
infections at time $t$.
We show that several such forecasting problems are computationally
intractable even when the specified time horizon $t$ is as small as 2.
We also present a result that shows the difficulty of obtaining
approximate solutions to these forecasting problems.
We demonstrate that our complexity results are tight by 
showing that the problems
are efficiently solvable for appropriate lower values of $t$.
Our hardness results also hold for realistic social networks
(e.g. power-law networks, small world networks).
We also provide efficient randomized approximation
algorithms for some forecasting problems.
Further, we extend our results to many other epidemiologically
relevant measures and to several other epidemic models.

\bigskip\bigskip

\noindent
\textbf{Note:}~ This report consists of two parts. 
\begin{itemize}
\item
Part~1, an abbreviated version of this report,
presents an overview of the problems considered, statements of results and 
their implications.

\item
Part~2 contains formal statements of problems and results as well 
as detailed proofs.
\end{itemize}

\smallskip
\noindent
To ensure that both parts are reasonably self-contained,
some of the text in Part~1 also appears in Part~2.


%%\begin{center}
%%\fbox{\fbox{\Large{\textbf{Please Do Not Distribute}}}}
%%\end{center}

\iffalse
%%%%%%%
\bigskip\bigskip


\noindent
\textbf{Acknowledgments:}~ 
This work has been partially supported by
DTRA Grant HDTRA1-11-1-0016 and 
DTRA CNIMS Contract HDTRA1-11-D-0016-0010.
%%%%%%
\fi

\setcounter{page}{0}
\clearpage
\thispagestyle{empty}

\vspace*{4in}

\begin{center}
\fbox{\Large{\textbf{Part 1:~ Overview of Problems and Results}}}
\end{center}

\setcounter{page}{0}
\clearpage

%%% New commands for the three state values.
\newcommand{\sstate}{\mbox{$\mathbb{S}$}}
\newcommand{\istate}{\mbox{$\mathbb{I}$}}
\newcommand{\rstate}{\mbox{$\mathbb{R}$}}

%% New command for the set with the above three values.
\newcommand{\bset}{\mbox{\{\sstate, \istate, \rstate\}}} 

\baselineskip=1.05\normalbaselineskip

\input{prob_names}  %%% Contains macros for problem names,

\input{sec_00}   %% Overview of problems and results (Part 1)

\clearpage
\setcounter{page}{0}
\thispagestyle{empty}

\vspace*{4in}

\begin{center}
\fbox{\Large{\textbf{Part 2:~ Details Regarding Problem Formulations and Results}}}
\end{center}

\setcounter{page}{0}
\setcounter{section}{0}
\setcounter{table}{0}
\clearpage
\baselineskip=1.29\normalbaselineskip

\input{sec_01}   %% Introduction.
\input{sec_02}   %% SIR a SSyDS -- Model definitions.
\input{sec_03}   %% Problem definitions and summary of results

\input{sec_04}   %% Intractability results for Forecasting Problems:
                 %% General networks.

\input{sec_05}   %% Additional Intractability results for 
                 %% Forecasting Problems:
                 %% More realistic social networks.

\input{sec_06}   %% Efficiently solvable or approximable versions
                 %% of Forecasting Problems.

\input{sec_07}   %% Extensions to other epidemic models.

\input{experiments}

\input{sec_08}   %% Conclusions, etc.

%%% Bibliography.
\clearpage
\baselineskip=\normalbaselineskip

%%\bibliographystyle{plain}
\bibliographystyle{abbrv}
\bibliography{sir_references.bib,epi-refs.bib}

\clearpage

\baselineskip=1.3\normalbaselineskip
\input{appendix}

\end{document}
