\noindent
{\Large\textbf{Background and Motivation}}

\epigraph{``Prediction is very difficult, especially if it's about the future".}
{--- \textup{Niels Bohr~ (1885--1962)}}  



\medskip
As large unexpected disease outbreaks are likely to have 
devastating economic consequences, there is increasing interest in
the development of systems that can provide early warnings regarding
epidemics.
This is borne out by the large number of epidemic forecasting challenges 
issued by various agencies; examples include 
the ``CHIKV Challenge'' by 
DARPA\footnote{\url{http://www.darpa.mil/news-events/2015-05-27}},
``Predict the Influenza Season Challenge'' by 
CDC\footnote{\url{https://www.federalregister.gov/documents/2013/11/25/2013-28198/announcement-of-requirements-and-registration-for-the-predict-the-influenza-season-challenge}},
the Dengue forecasting challenge by 
NOAA\footnote{\url{http://dengueforecasting.noaa.gov/}},
the NSF/NIH Ebola forecasting challenge and 
the IARPA Flu challenge \cite{Muthaiah_etal_2016}.
%%% Commented out since the URLs couldn't be found. -- Ravi
%% \textcolor{red}{URLs for these?}. 
%%%
There has also been a burst of activity on attempting to forecast different kinds of
phenomena such as epidemic outbreaks
\cite{nsoesie2013systematic,Nishiura:2011,Ohkusa:2011,Hall:2007,
Vespignani-BMC-2012, chakraborty:sdm14, Scarpino-Petri-2017},
cascades in social media and civil unrest
(e.g., \cite{Martin_etal_2016, Krishnan_etal_2016, Korkmaz_etal_2016,
Ramakrishnan:2014:BNE:2623330.2623373}).
The difficulty of forecasting geophysical phenomena 
such as earthquakes has also been noted in several papers
(e.g., \cite{Geller-1997,Geller-etal-1997}).
Many articles in a recent issue of \emph{Science} 
(Volume: 355, Issue: 6324, February 2017) point out
the difficulty of accurately forecasting the behaviors of
complex social systems \cite{Athey-2017,Bohannon-2017,Cederman-etal-2017,
Clauset-etal-2017,Hofman-etal-2017,Jasny-etal-2017,Kennedy-etal-2017,
Subrahmanian-etal-2017,Tetlock-etal-2017}.

\medskip
Despite a lot of work, epidemic forecasting remains poorly understood.  
A case in point is Google Flu Trends (GFT) 
\cite{ginsberg:nature08:flu-search-engine}. 
In its initial years, GFT
produced very good forecasts of flu incidence rates,
just based on search query results. 
However, the forecast accuracy decreased over time
(e.g., overestimates of the A/H3N2 epidemic
\cite{lazer:science14, olson:ploscb13}).
Since the spread of flu-like diseases is a complex stochastic process 
which depends on many time-varying factors \cite{Drake-2005, Drake-2006},
one cannot expect accurate long-term predictions
from a model such as GFT that relies on just one form of data.
It is pertinent to note that the data generated by GFT is being used at several 
institutions in developing new models for disease dynamics;
however, several fundamental questions regarding 
what can be forecast remain open.

\medskip
Many researchers have observed the need for carrying out
a systematic study of the various issues associated with forecasting 
contagion dynamics in networked systems.
Some of these observations are summarized below.

\begin{itemize}
\item
Drake \cite{Drake-2005, Drake-2006} observes that
while systems may provide good forecasts 
of some epidemic  measures (e.g., timing),
there are fundamental limits on the effectiveness
of such systems in forecasting other measures (e.g., final epidemic size).
Several important reasons (such as the nature of stochastic 
disease propagation models and high sensitivity of disease parameters
to changes in the environment)
are articulated in \cite{Drake-2005,Drake-2006}
to explain such limits. 

\item
Cheng, Adamic, Dow, Kleinberg and Leskovec \cite{Cheng_etal_2016}
address the question of predicting whether a cascade will continue
to grow in a social network.  
In particular, they focus on identifying the features of a cascade
that can help in predicting the future course of the cascade.
They remark that ``a robust way to formulate the problem of cascade
prediction remains an open problem".

\item
Martin, Hofman, Sharma, Anderson and Watts \cite{Martin_etal_2016}
examine the limits of predictability
in complex social systems; they observe that even a small degree of
uncertainty can limit predictability.
Further, they raise the question of whether a phenomenon that
one is trying to predict is itself inherently unpredictable or 
the available data and techniques are inadequate to develop a
reliable prediction.
In another article, Hofman, Sharma and Watts \cite{Hofman-etal-2017}
mention that ``theoretical limits to the predictive accuracy of
complex systems must be better characterized".

\item 
Lazer, Kennedy, King and Vespignani \cite{lazer:science14} use 
Google Flu Trends as an example to caution against the use of social media 
and search information as a substitute for traditional data collection
and analysis methods to predict epidemic measures.
They highlight the need for systematically ``studying the evolution of the
socio-technical systems that are embedded in our society".

\item
Shaman et al. \cite{Shaman-etal-2012,Shaman-etal-2013}
use a differential equation based model along with new
data analysis techniques to predict the timing of the peak number of
infections for influenza outbreaks.
They demonstrate the usefulness of their framework in
predicting the peak timing for the 2012--2013 season.
They suggest that their methods can be made more robust 
when additional data is available. 

\item 
Biggerstaff et al. \cite{Biggerstaff-etal-2016} discuss results from 
the Centers for Disease Control and Prevention's (CDC's) ``Predict the
2013--2014 Influenza Season Challenge". 
They conclude that ``forecasting has become technically feasible, but
further efforts are needed to improve the accuracy so that policy makers can
reliably use those predictions".

\item
May \cite{May-2006} points out that care must be exercised in
drawing conclusions about the dynamics of ecological systems when
there is uncertainty in the structure and parameters of the 
underlying network model. 
\end{itemize}

Beckage, Gross and Kauffman \cite{beckage:ecosphere11} 
point to another another challenge
for forecasting, which they refer to as ``computational irreducibility''---this 
is the property of systems where
the dynamics cannot be predicted without observing the evolution of the system.
However, this can be difficult if determining the properties associated with
the evolution of a system is itself computationally demanding---this is referred to as 
\emph{computational intractability}
\cite{GJ-1979}, and has been a cornerstone of modern theoretical computer science.
Indeed, many researchers (e.g., Buss, Papadimitriou and Tsitsiklis \cite{Buss-etal-1991},
Moore \cite{Moore-1990}, Wolfram \cite{Wolfram-1985, Wolfram-1986})
have observed that computational intractability results for discrete
dynamical systems provide an indication of the unpredictability 
(or ``chaotic" behavior) of such systems.
The theory of computational intractability has helped 
to classify computational problems into a number of 
classes, such as \textbf{P}, \cnp, \cnump, etc. 
(see e.g., \cite{GJ-1979} for an introduction to this topic).
Informally, problems that are \cnp-hard or \cnump-hard are unlikely to 
have efficient algorithms, that is, algorithms
which run in time that is a polynomial function of the input size.
Computationally intractable problems arise in various fields 
(e.g., Mathematics, Physics, Biology, Social Science, Computer Science, 
Operations Research).  
In the context of forecasting,
some computational intractability results that arise in 
testing weather forecasts are presented in \cite{Fortnow-etal-2009};
however, the model used in their work is different our network-based
epidemic model. 

\medskip
In this paper, we show that many fundamental problems related to 
short-term predictions of epidemic properties
are computationally intractable  
even when all the model parameters are known and 
assumed to be insensitive to changes in the environment.
Using the observations in \cite{Buss-etal-1991,Wolfram-1985,Wolfram-1986,Moore-1990},
these results are indicative of the unpredictability of epidemic dynamics
in networked systems.
Thus, our results bring out a fundamental difficulty of 
predicting disease parameters. 
%%A detailed discussion on the implications of our results is given later.
%%\cite{Drake-2005,Drake-2006,beckage:ecosphere11} 
%% even when all the model
%%parameters are known and assumed to be insensitive to changes
%%in the environment.

Many computational models for the study of
epidemics have been proposed in the literature (see e.g., 
\cite{Brauer-2008,marathe:cacm13,Britton-2009,kermack+m:SIR,epi-Bailey,
epi-Hethcote}).
%% These models are useful to public health officials in both 
%% predicting the spread of an epidemic and in developing measures
%% to contain the epidemic.
Our results hold under the well known
Susceptible-Infected-Recovered (or SIR) model 
initially proposed in \cite{kermack+m:SIR}, 
as well as under a broad class of 
related models, such as SI, SIS and probabilistic threshold 
(complex contagion) models. 
We note that our results hold for networked populations, and not
in homogeneous mixing of populations, where the SIR and similar
models were originally studied.
%%A precise description of the SIR model on networks is
%%presented in Section~\ref{sec:sir_dyn_system}.
%%As observed by a number of researchers 
%%(e.g., \cite{marathe:cacm13,SD-2012,Bansal_etal_2007,eubank:nature04,
%%newman:spread02,Cauchemez15022011}), contact patterns arising
%%in practice do not satisfy such homogeneity assumptions.

\bigskip

%\clearpage

\noindent
{\Large\textbf{Problems Studied}}

\medskip
\noindent
We introduce a number of short-term forecasting
problems under the SIR model. 
These problems can also be posed under other epidemic
models mentioned above.
An example of a problem (denoted by \tNewInfs) under this model 
is the following: given
a social network, the set of initially infected nodes, a subset $S$
of nodes, an integer $q \leq |S|$ and a time value $t$, find the probability 
that at least $q$ nodes of $S$ get infected \emph{at} time $t$.
Another problem (denoted by \tPeak) asks for the probability that
the number of new infections reaches a peak at time $t$.
Several such problems can be formulated and
Table~\ref{init:tab:prob_def} gives\footnote{All the tables appear on
pages 5, 6 and 7.} concise descriptions of the main problems
studied in our work. 

\bigskip

\noindent
{\Large\textbf{Summary of Results}}

\medskip
\noindent
%% The SIR model has been studied extensively. Many results
%% regarding the conditions under which the disease will spread rapidly and
%% techniques to control the spread have been reported (e.g., 
%% see \cite{epi-Hethcote,Britton-2009,Brauer-2008,marathe:cacm13} and the
%% references cited therein).
To our knowledge, only a few references \cite{LZ-2009,PS-2012,SD-2012}
have addressed the computational complexity issues related to forecasting 
under the SIR or other epidemic models\footnote{
We provide a summary of these results in Section~\ref{sse:related_work}
(in Part~2 of this report).}.
However, none of these references addresses short-term forecasting problems.
Our contributions can be summarized as follows.

\begin{enumerate}
\item For general graphs, we establish computational intractability results for
the four problems, namely \\
\tNewInfs,~ \tTotInfs,~ \tVuls{} and \tTotVuls,\\
even when the time horizon $t$ is as small as 2.
We also prove a result that provides an indication of the difficulty
of obtaining approximate solutions to any of these problems
for any $t \geq 2$.
In addition, we demonstrate the tightness of our intractability results 
by showing that the
four problems are efficiently solvable for $t = 1$.
Further, we show that the \tPeak{} problem is computationally
intractable even for $t = 1$.
We also present a randomized approximation scheme for the problem
\tTotVuls, for any fixed $t$ and any set $S$ of fixed size.
Table~\ref{init:tab:gen_results} provides precise statements of
our results for general graphs.

\item We extend the above intractability results to more
realistic social networks (e.g., networks with low diameter and
high clustering coefficient, power-law networks) and to networks
in which all the edges have the same transmission probability.
These results are shown in Table~\ref{init:tab:realistic_results}.

\item We also present extensions of our results to prove the intractability
of forecasting various epidemic measures introduced in \cite{TC+2016}.
These results are summarized in Table~\ref{init:tab:new_measures_results}.

\item Further, we extend our intractability results to
three other epidemic models, namely SI, SIS and PT
(probabilistic threshold).
These results are summarized in Table~\ref{init:tab:results_for_models}.
\end{enumerate}
In a companion paper \cite{Rosenkrantz_etal_2016}, we show that
many of the forecasting problems under the SIR model can be solved efficiently
when the treewidth \cite{Bod93} of the underlying graph is bounded.

%\bigskip


%\bigskip

\clearpage

\noindent
{\Large\textbf{Implications}}

\medskip\smallskip

\noindent
\textbf{1.~Unpredictability of epidemic dynamics sets in at a very early time.} \ 
Our results show that from a computational complexity point of view, 
epidemic dynamics becomes unpredictable even when the 
time horizon is as small as 2.
This is in contrast to previous complexity results on the unpredictability
of dynamical systems which needed an exponential number of time steps for
deterministic systems (e.g., PSPACE-hardness of reachability problems
for dynamical systems shown in \cite{BH+06}) and a polynomial number of time
steps for probabilistic systems (e.g., PSPACE-hardness of simulating 
quadratic dynamical systems shown in \cite{Arora-etal-1994}).

\medskip
\noindent
\textbf{2.~Unpredictability holds for both macroscopic and individual properties.} \ 
Our computational intractability results hold for predicting 
macroscopic properties (e.g., finding the probability that there will be at least
$q$ new infections at any time $t \geq 2$) as well as individual properties
(e.g., finding the probability that a specific node is infected at time $t = 3$
under the probabilistic threshold model) over a short time horizon.

\medskip
\noindent
\textbf{3.~Unpredictability holds even when prior knowledge regarding
the system behavior is available.} \ 
Most of our computational intractability results hold for any time value $t \geq 2$.
An examination of the proofs of these results shows that
from a worst-case stand point, the unpredictability results do not change
even if the behavior of the system is known for most previous time steps. 

\medskip
\noindent
\textbf{4.~The unpredictability results are pervasive.} \
Our unpredictability results hold for a variety of problems 
and contagion propagation models. 
The problem variants include predicting the number of new infections
at a certain time, the total number of infections up to a certain time,
the time and size of the peak number of new infections, etc.
The contagion models studied include SIR, SI, SIS and probabilistic
threshold. 
Furthermore, the complexity results hold for
several classes of networks such as power-law networks and small world
networks (e.g., complete graphs whose diameter is 1).

\medskip
\noindent
\textbf{5.~One random parameter is adequate for unpredictability.} \ 
Our complexity results rely on exactly one source of randomness, namely
the transmission probability. 
The results hold even when the remaining model parameters 
(e.g., the network structure, 
incubation time, infectious period, initially infected node)
are deterministic and do not vary over time.  

\medskip
\noindent
\textbf{6.~Provable predictability results for certain special cases.} \
For certain special forecasting problems, our results provide 
provably good approximations when the time horizon is fixed.
Such special cases include computing the expected number of infections,  
the probability of a node being infected by a certain time and the
probability that the total number of infections exceeds a specified 
count by a certain time.  
%%The last results are
%%interesting and are based on using an importance sampling technique.

\smallskip
In summary, our results show that one cannot expect to find
provably correct and efficient algorithms for predicting 
epidemic dynamics on networks.  
The results do not rule out heuristics that work well in practice;
they suggest that good algorithms that work across a range of inputs 
should exploit special properties of problem instances.

%%%%%%%%%% All the tables are here. %%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[tbh]
\bigskip
\begin{center}
\begin{tabular}{|p{1.6in}|p{3.9in}|} \hline
\multicolumn{1}{|c|}{\textbf{Problem Name}} &
\multicolumn{1}{|c|}{\textbf{Description}} \\ \hline\hline
\tNewInfs{} & {Compute the probability that the number of new infections
              within a given subset $S$ of nodes \emph{at} time $t$
           is at least $q$.} \\ \hline
\tTotInfs{} & {Compute the probability that the total number of
              infections within a given subset $S$ of nodes \emph{by} time $t$ is
              at least $q$.} \\ \hline
\tVuls{} & {Compute the vulnerability of the nodes in subset $S$ \emph{at}
            time $t$ (i.e., the probability that \emph{all} the nodes
                         in the subset $S$ get infected
                         \emph{at} time $t$).} \\ \hline
\tTotVuls{}   & {Compute the vulnerability of the nodes
                             in $S$ \emph{by} time $t$
                 (i.e., the probability that \emph{all} the nodes in $S$ get infected
                             \emph{by} time $t$.)} \\ \hline
\tPeak{}   & {Compute the probability that the number of new infections in
              the network reaches a peak \emph{at} time $t$.} \\ \hline\hline
\end{tabular}
\end{center}
\caption{Concise descriptions of the forecasting
problems considered in the paper.
Problems \tVuls{} and \tTotVuls{} are special versions 
of \tNewInfs{} and \tTotInfs{} respectively with $q = |S|$.
When $S = V$ (the set of all nodes in the network), we denote
the first four problems by \tNewInfv, \tTotInfv,~ \tVulv{}~ and \tTotVulv{}~
respectively.
}
\label{init:tab:prob_def}
\end{table}


\begin{table}[tbh]
%%\begin{table}
\begin{center}
\begin{tabular}{|p{1.6in}|p{4.15in}|} \hline
\multicolumn{1}{|c|}{\textbf{Problem(s)}} & 
\multicolumn{1}{|c|}{\textbf{Result(s)}} \\ \hline\hline   
{\tNewInfs, \newline
\tTotInfs, \newline 
\tVuls{}~and \newline
\tTotVuls{}}
              & {(a)~ \cnump-hard for any~ $t \geq 2$ ~(Part~(1) of 
                     Theorem~\ref{thm:gen_hardness}). 

                \smallskip
                (b)~ For any $t \geq 2$, there is an $\epsilon > 0$ such
                     that unless \textbf{P} = \cnp, the quantity 
                     $\log{(2^n\,p^*)}$ cannot be efficiently approximated to
                     within the factor $n^{\epsilon}$, where $p^*$ is the solution
                     value and $n$ is the maximum number of nodes that can
                     get infected at $t = 1$. (Part~(2) of \newline
                     Theorem~\ref{thm:gen_hardness}). %\newline
               } \\ \hline
{\tPeak{}}  & {\cnump-hard for  any $t \geq 1$~ (Part~(3) of
               Theorem~\ref{thm:gen_hardness}).
              } \\ \hline\hline
{\tNewInfs, \newline
\tTotInfs, \newline
\tVuls{}~and \newline
\tTotVuls{}}
            & {Efficiently solvable for~ $t = 1$~
                     (Part~(1) of Corollary~\ref{cor:gen_easy_time_1}).} \\ \hline
{Compute the expected number of new infections at time $t$} 
            & {Efficiently solvable for~ $t = 1$~ and~ $t = 2$~
                     (Corollary~\ref{cor:expected_number_infected_time_2}).} \\ \hline
{\tTotVuls{}}  & {Randomized approximation scheme for 
                 any fixed $t$ and any set $S$ of fixed size~ 
                 (Theorem \ref{thm:totvul_approx}).
              } \\ \hline\hline
\end{tabular}
\end{center}
\caption{Results for forecasting problems over general graphs.
(The theorems and corollaries mentioned in the above table
appear in Part~2 of this report.)
\bigskip} 
\label{init:tab:gen_results}
\end{table}

\begin{table}[tbh]
%%\begin{table}
\begin{center}
\begin{tabular}{|p{1.6in}|p{3.9in}|} \hline
\multicolumn{1}{|c|}{\textbf{Problem}} & 
\multicolumn{1}{|c|}{\textbf{Result(s)}} \\ \hline\hline   
{\TwoNewInfv{}} & {\cnump-hard even when both the diameter 
                      and the average clustering coefficient are 1
                       (Part~(1) of Theorem~\ref{thm:dia_cc}).
                 } \\ \hline
{\TwoTotInfv{}} & {\cnump-hard even when both the diameter 
                       and the average clustering coefficient are 1
                       (Part~(2) of Theorem~\ref{thm:dia_cc}).
                 } \\ \hline
{\tNewInfv{}}   & {(a)~ \cnump-hard for any $t \geq 2$ 
                      even when all edge probability values \newline
                      are less than 1 ~(Part~(1) of 
                      Theorem~\ref{thm:gen_hardness_prob_not_one}). \newline
                  (b)~ \cnump-hard for any $t \geq 3$ even when 
                       all edge probability values \newline
                       are equal to $0.5$~
                       (Theorem~\ref{thm:gen_hardness_prob_half}). \newline
                  (c)~ \cnump-hard for any $t \geq 2$ even 
                       for power-law graphs\newline 
                       (Theorem~\ref{thm:hardness_power_law}).
                 } \\ \hline
{\tTotInfv{}}   & {(a)~ \cnump-hard for any $t \geq 2$ 
                       even when all edge probability values \newline
                       are less than 1 ~(Part~(2) of
                       Theorem~\ref{thm:gen_hardness_prob_not_one}). \newline
                  (b) \cnump-hard for any $t \geq 3$ even when 
                      all edge probability values \newline 
                      are equal to $0.5$~
                      (Theorem~\ref{thm:gen_hardness_prob_half}).
                 } \\ \hline\hline
\end{tabular}
\end{center}
\caption{Extensions of intractability results to more realistic networks.
(Proofs of the  theorems mentioned in the above table
appear in Part~2 of this report.)}
\label{init:tab:realistic_results}
\end{table}

\begin{table}[tbh]
\begin{center}
\begin{tabular}{|c|l|}\hline
\multicolumn{1}{|c|}{\textbf{Problem}} &
\multicolumn{1}{|c|}{\textbf{Result}} \\ \hline\hline
Peak Value Forecasting (PVF) & \cnump-hard \\ \hline
Take-Off Value Forecasting (TOVF) & \cnump-hard \\ \hline
Take-Off Time Forecasting (TOTF) & \cnump-hard for any $t \geq 1$ \\ \hline
Intensity Duration Forecasting (IDF) & \cnump-hard for any duration $\tau \geq 2$ \\ \hline
Epidemic Speed Forecasting (ESF) & \cnump-hard \\ \hline
Total Attack Rate Forecasting (TARF) & \cnump-hard for any $t \geq 2$ \\ \hline\hline
\end{tabular}
\end{center}
\caption{Results for Forecasting Problems for Various Epidemiological
Measures defined in \cite{TC+2016}.
(The definitions of the problems mentioned in this table
are given in Part~2 (Section~\ref{sse:ext_measures}) of this report.
Formal statements of the results mentioned
in the table are given in Theorem~\ref{thm:hardness_new_measures} 
(Section~\ref{sse:ext_measures}).)}
\label{init:tab:new_measures_results}
\end{table}

\begin{table}[tbh]
\begin{center}
\begin{tabular}{|p{1.5in}|p{4in}|}\hline
\multicolumn{1}{|c|}{\textbf{Model}} &
\multicolumn{1}{|c|}{\textbf{Results}} \\ \hline\hline
{SI Model} & {(a) \tNewInfs, \tTotInfs{} and \tPeak{} are 
             \cnump-hard for any $t \geq 2$. \smallskip  \newline
           (b) There is a randomized approximation scheme for \newline
           \tTotVuls{} when $t$ and $|S|$ are fixed.
           (The quantity being approximated is the probability that each
            node in $S$ gets infected at least once by time $t$.)} \\ \hline
SIS Model & {(a) \tNewInfs, \tTotInfs{} and \tPeak{}
             are \cnump-hard for any $t \geq 2$.\smallskip  \newline
           (b) There is a randomized approximation scheme for \newline
           \tTotVuls{} when $t$ and $|S|$ are fixed.} \\ \hline
Probabilistic Threshold (PT) Model & {(a) \tNewInfs, \tTotInfs{} and \tPeak{} are
            \cnump-hard for any $t \geq 2$. \smallskip  \newline
           (b)  \tVuls{} is \cnump-hard for any fixed $t \geq 3$ even
              when $|S| = 1$;~ however, the problem is
           efficiently solvable for any $S$ for $t = 1$ and $t = 2$.} \\ \hline\hline
\end{tabular}
\end{center}
\caption{Summary of Results for Forecasting Problems Under Various Models
(Proofs of these results appear in Section~\ref{sse:ext_models} of Part~2.)}
\label{init:tab:results_for_models}
\end{table}

%%%%%%%%%% End of TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%% From here on, all subsequent text has been %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% commented out.                             %%%%%%%%%%%%%

\iffalse
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\medskip
\noindent
A direct consequence of our results is that several basic forecasting 
problems that arise in the context of contagion dynamics cannot 
be solved efficiently, unless widely believed hypotheses in complexity
theory turn out to be false.
These intractability results hold even when 
the time horizon is as small as 2.
Further, one cannot even hope for efficient
algorithms that produce provably approximate solutions to the 
forecasting problems.
Recall that the forecasting problems
remain computationally intractable for
realistic social networks (e.g., small world networks, 
power law networks).
As a consequence, algorithms that
produce exact (or provably approximate) solutions for
large social networks that arise in practice
are unlikely to run in a reasonable amount of time.

\medskip
Other researchers 
(e.g., \cite{Drake-2005,Drake-2006, Martin_etal_2016,May-2006,beckage:ecosphere11})
have attributed the difficulty of making accurate 
forecasts to the uncertainty in the values of model parameters.
While parameter uncertainty is indeed a source of difficulty,
our results point to a fundamental computational limitation on
accurate (or approximate) forecasts, even when the 
model parameters are known exactly and are 
insensitive to changes in the environment.
Thus, our results supplement the reasons discussed in 
\cite{Drake-2005,Drake-2006, Martin_etal_2016,May-2006,beckage:ecosphere11}
by bringing out another intrinsic difficulty associated with
predicting parameters of contagion dynamics.

\medskip
Our randomized approximation schemes show that for forecasting
problems involving small subpopulations and a small time horizon,
one can obtain provably good approximations efficiently. 
These results suggest that restricting to a suitable 
subpopulation is a useful step in obtaining reliable and nearly optimal
forecasts related to contagion dynamics.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi

\iffalse
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bigskip

\noindent
{\Large\textbf{Remarks About the Techniques Used}}

\medskip
\noindent
The \cnump-hardness results and the difficulty of approximating
the solution value shown in Table~\ref{init:tab:gen_results}
are obtained through reductions from the problem of counting
the number of satisfying assignments to the Monotone 2SAT problem
studied in \cite{Vad-2001,Zu-1996}.
We prove the efficient solvability of the problems for $t = 1$ by 
developing a dynamic programming algorithm for a more
general problem that encompasses \tNewInfs,
\tTotInfs, \tVuls{} and \tTotVuls.
The randomized approximation scheme for \tTotVuls{}
(when $t$ and $|S|$ are fixed) is based on a reduction to the
problem of counting the number of satisfying assignments to
a Boolean formula in disjunctive normal form (DNF). 
In establishing some of the results for 
realistic (e.g., small world) networks 
(see Table~\ref{init:tab:realistic_results}),
we use FKG inequalities from probability theory \cite{FKG_1971}.
The results presented in Tables~\ref{init:tab:new_measures_results}
and \ref{init:tab:results_for_models} use appropriate modifications
of the constructions used to prove the results mentioned in 
Table~\ref{init:tab:gen_results}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi

